AtmoSwing (v1.4): Analog Technique Model for Statistical weather forecastING

### 2016-08-17 AtmoSwing - Geoscientific Model Development/template.tex
Changing "analogue" for "analog".

\title{AtmoSwing (v1.4): Analogue Technique Model for Statistical weather forecastING}

@@

The Analogue Method allows forecasting local meteorological variables of interest (predictand), such as the daily precipitation, on the basis of synoptic variables (predictors). It relies on the fact that similar meteorological influences are likely to result in similar local effects. This statistical relationship is thus based on archives of observed data, and used in order to operationally forecast the coming days, or to evaluate future conditions under a changing climate.

AtmoSwing is an open source software that implements the Analogue Method in a very flexible way, so that different variants can be handled dynamically, by parameterization through XML files. It is made of 3 tools: the Forecaster to perform operational forecasts, the Viewer to display the results, and the Optimizer to establish the relationship between the predictand and predictors.

@@

Usually, the predictand values could be derived in a rather physical-dynamic approach  by modeling the chain of processes linking the predictors to the predictand. The processes involved range from large scale dynamical states of the atmosphere down to very small scale microphysical processes, requiring models that are extremely complex, data-demanding, and time consuming. Conversely, given an appropriate set of predictors archives, enough situations analogous to a target one could be found so that reasonable values will be obtained for the predictand, at a reasonable coding and computing-time costs. This is particularly true for a predictand much required in hydrometeorological applications, namely the precipitation amount over a given domain and time duration. Incidentally, the forecast will be proposed as a statistical distribution based on the values taken by the predictand in the set of analogues selected, unless one considers only the single best analogue, which may not prove the most efficient \citep{Bontron2005}.

In addition to precipitation forecasting, the Analogue Method (AM), or an equivalent, was also used for short to medium term forecasting of daily temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983}, wind \citep{Gordon1987}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}. \citet{Guilbaud1997} performed a literature review about the use of the AM in long-term forecasting and identified operational applications for monthly forecasts in many countries, including Canada \citep{Shabbar1986},  Hungary \citep{Toth1989}, the Netherlands \citep{Nap1981}, and England \citep{Murray1974}, as well as seasonal forecasts: \citet{Barnett1978}, \citet{Bergen1982} and \citet{Livezey1988}.

@@

\section{The Analogue Method}
\label{sec:method}

@@

The AM is based on the principle that two similar synoptic situations may produce similar local effects \citep{Lorenz1956}. The perfect analogy does not exist, but sufficiently similar situations leading to similar effects can be identified. Thus, two states of the atmosphere that are alike are called analogues \citep{Lorenz1969}. To be relevant, this analogy must be selected by optimizing the following elements:

\begin{itemize}
	\item The analogy variable (the predictor) must contain synoptic information having indirect dependency with the local weather and the target predictand (often the precipitation).
	\item The pressure level, or height, at which the analogy variable is considered.
	\item The spatial window is the domain on which the analogy variable is considered. The ideal size of this area is the one that maximizes useful information.
	\item The temporal window is the hour in the day, or a set of different hours, at which the analogy variables are considered.
	\item The preselection, which is often a period of the year in which similar situations are being sought. Thus, only situations with a similar distribution of solar energy are compared \citep{Lorenz1969}.
	\item The analogy criterion, needed to compare the variables on the chosen spatial and temporal windows, is a score used for ranking past situations according to their degree of similarity with the target situation.
	\item The ideal number of analogue situations is the best compromise in order to take into account local variability and to maximize useful synoptic information \citep{Bontron2004}.
\end{itemize}

Because of the chaotic nature of the atmosphere, two analogue situations quickly diverge over time \citep{Lorenz1969}. Thus, the AM has strong limitations regarding the temporal extrapolation \citep{Bontron2004}. Numerical models being more capable of simulating the dynamic evolution of the atmosphere, the temporal extrapolation of the synoptic variables is left to them. The search for analogy aims thus at connecting the forecasted synoptic situation with a local predictand (temperature, precipitation, etc), which is more difficult to simulate by numerical models.

@@

	The $N_{1}$ dates with the lowest values of S1 are considered as analogues, in terms of the general circulation, to the target day. The number of analogues, $N_{1}$, is a parameter to calibrate.

	\item Subsequent level(s) of analogy: beyond the similarity of airflows, one may look for similarities in other variables of interest, such as moisture variables. Therefore, the $N_{1}$ analogues are subsampled once again on the basis of another variable in order to obtain a lower number of analogue dates, $N_{2}$. The S1 criterion is usually not relevant for other predictors that the atmospheric circulation. Classic criteria represent absolute distances: Mean Absolute Error (MAE) and Root Mean Squared Error (RMSE), the latest being most often used.

	This process can be repeated, by subsampling a decreasing number of analogues, $N_{i}$, according to various meteorological variables.

@@

The basis of the method was introduced by \citet{Lorenz1969}, who considered the analogy of the geopotential field on different pressure levels. In addition, the analogue situations were to belong to the same period of the year to be comparable in terms of distribution of solar energy. This preselection was later implemented as a moving selection of more or less two months around the target date.

@@

Every variable of the NCEP/NCAR reanalysis was systematically assessed for each of the identified subregions of the study area (Sect. \ref{sec:case_study}). The present assessment considers every subregion separately and optimizes the spatial window and the number of analogues for every combination of variables, pressure levels, times of observation, and subregions (using the sequential calibration tool implemented in AtmoSwing, see Sect. \ref{sec:atmoswing-calibration}). Even though the best predictors may vary from one place to another due to different meteorological influences (see Fig. \ref{figure:variable_exploration_Chablais} for the 30 best predictors for the Chablais region, and Fig. \ref{figure:variable_exploration_SE_crests} for the Southeast ridges), 2 main conclusions can be drawn from the NCEP/NCAR reanalysis dataset for the area of interest, that meet those from \citet{Bontron2004}: (1) the variables describing the atmospheric circulation (pressure fields or geopotential heights) are the ones with the highest prediction capacity (the precipitation rate is even rarely selected in the 30 best predictors), and (2) these are better when compared by means of the S1 criteria (asterisk in Fig. \ref{figure:variable_exploration_Chablais} and \ref{figure:variable_exploration_SE_crests}) instead of the RMSE. The reason for this latter point is the better representation of the atmospheric circulation information when considering the shape of a pressure field rather than its values.

\citet{Horton2012} identified that the analogy on the atmospheric circulation as defined by \citet{Bontron2004} may not be comprehensive in an alpine environment, and that considering 4 atmospheric levels (1000, 850, 700, and 500~hPa at 12 or 24~h) instead of 2 could improve the performance (Table \ref{table:R6} -- the number of analogues $N_{1}$ was not provided globally, as it was calibrated for every subregion). The change in performance scores is illustrated in Fig. \ref{figure:scores_1} for the different subregions of the study area. An equivalent structure was also optimized by means of generic algorithms (PC-4Zgo), which is the topic of another paper \citep{Horton2016b}. As these new parametrizations provide performance gains for alpine catchments under varying meteorological influences (Fig. \ref{figure:scores_1}), they are recommended for operational use.

@@

\citet{Horton2012} also integrated an analogy on the moisture flux, after the circulation analogy on the 4 atmospheric levels (Table \ref{table:R7}). The number of analogues ($N_{1}$, $N_{2}$) were calibrated for every subregion. Similarly, an equivalent structure was optimized \citep[PC-4Zgo-2MIgo,][]{Horton2016b}. The change in the performance scores, for the method with 2 levels of analogy, is illustrated in Fig. \ref{figure:scores_2} for the study area. As previously, we would recommend using these new parametrizations for alpine catchments.

@@

\citet{Hamill2006} used an analogy based approach on the GFS reforecasts in order to correct systematic errors in the ensemble forecasts of temperature and precipitation. Indeed, the statistical approach pointed at some bias in estimates of the numerical model, which may have been adjusted by taking into account the intrinsic local climatology from the AM. Moreover, the under-dispersion of the ensemble forecast from the numerical model has been corrected using analogues \citep{Hamill2006}. Correction of ensemble forecast under-dispersion by means of the AM is also used operationally at EDF (\'{E}lectricit\'{e} de France).

@@

Another limitation is the fact that extreme events may be under-represented in the considered sample of analogue dates. Indeed, in a limited weather archive, events with high return periods are not so numerous. Their number is certainly lower than the standard number of analogues considered, which can introduce a bias in the forecast. There are however techniques to correct this bias \citep[see][]{Marty2010}.

@@

The source code in under version control (with Mercurial), and is open source (on Bitbucket, www.atmoswing.org). Developments have been made with a test driven development (TDD) approach. A collection of more than 500 tests are frequently evaluated and completed during development, in order to verify that code changes do not introduce regression. Every analogy criterion, prediction score, searching and sorting functions, data manipulation, and so on, is tested. The integration tests specific to the AM rely on the results of another analogue sorting software developed at the LTHE \textendash INPG university of Grenoble. They ensure that the results of AtmoSwing are exactly equivalent to their model, given the same parameters and data. These tests are part of the checks carried out regularly.

@@

An AtmoSwing great strength is that it is designed to process analogy forecasting in a modular fashion. The structure of the AM (number of analogy levels, number of predictors) is built dynamically (Fig. \ref{figure:flowchart_modules_atmoswing}), and nothing is fixed a priori. The software then performs successively as many analogy levels as the user specified, with all the predictors he wants. Each level of analogy processed results in an object containing target dates, analogue dates, values of the analogy criteria, and other data. This object can be saved in a NetCDF file and/or can be injected into a new analogy level. The whole structure of the AM is defined in an XML file generated by the user. Even the time step of the method (6 or 24~hours for example) is a dynamic parameter.

@@

Although processing an analogue forecast for a given target date is fast, when optimizing the AM, forecasts are re-processed over long periods of several decades, which may become very time consuming. Thus, great effort has been made in order to reduce the processing time to a minimum. Waste of time were identified with profiling tools and cut down at different levels. The software also use the linear algebra library Eigen 3 \citep{Guennebaud2010} for calculations on vectors and matrices, which resulted in time saving. Multi-threading is also implemented so that the search for analogue situations in the archive is distributed among the available threads.

The code optimization consited in reducing the most significant processing time identified by means of profiling tools. First, every identified redundancy in processing has been removed. Then, when looking for a certain date or data, the search looks first in the proximity where it is likely to be found instead of exploring an entire array. Similar data are not loaded twice, but shared pointers are used. Many other improvements allowed saving more time, for example the use of the quicksort method \citep{Hoare1962a} to sort the date vectors according to the values of analogy criteria. Different implementations were tested (many remain in the code as alternatives) in order to select the most efficient: for example, when storing analogue dates according to their criteria value, it is faster to insert them in a fixed-size array rather than storing them all and sorting the array subsequently.

The major performance gains were obtained by:
\vspace*{-2mm}
\begin{itemize}
	\setlength\itemsep{-2px}
	\item reduction of the array sizes used in the calculations,
	\item use of pointers (and therefore the decrease of data copies),
	\item better search for dates in temporal vectors starting from the previous index (reduction of the range of the array to explore),
	\item better management of the analogue vectors,
	\item preprocessed gradients on geopotential fields when using the S1 criteria,
	\item and finally the implementation of parallel calculations (multi-threading).
\end{itemize}

@@

The software first downloads the predictor describing the target situation, such as 1\textdegree\ GFS outputs \citep[Global Forecast System,][see Sect. \ref{sec:data}]{Kanamitsu1991,Kanamitsu1989}. It then interpolates linearly the gridded data to match the resolution of the archive (for example 2.5\textdegree\ in the case of the NCEP/NCAP reanalysis I). The analogues dates are next sought according to the selected parameterization, and the predictand data are associated with the corresponding dates. The results are finally saved in auto-describing NetCDF files. If required, a synthetic XML is generated for an easier integration on a web platform for example. Every step of the forecast, from the predictor downloading to the final results, is done in the software (and controlled through configuration), without use of external scripts (e.g. for data conversion).

When using a given parameterization of the AM, it has previously been optimized in a perfect prognosis context \citep[i.e. by optimizing on the reanalysis dataset; however, this does not take into account the uncertainty of the target situation when taken from operational forecasts of a Numerical Weather Prediction model,][]{Klein1963}. In operational forecasting, one should take into account the uncertainty of the predictor increasing with the lead time. A solution, presented in Sect. \ref{sec:optimization-operational}, consists in increasing the number of analogue situations to keep with the lead time. This technique is available in AtmoSwing, as the number of analogues can be specified for every lead time, making such approach easily accessible.

The command-line interface allows to start the process for the present date, a selected past date, or the last $x$ days (as long as the predictors are available on the provider server). When there is nothing to process (no new predictor data available), the execution just stops, leading to no loss of computing resources. The recommended use is thus to set up a cron task on a Linux server in order to trigger the forecast every 30 minutes. When using GFS outputs, this would provide 4 forecasts a day with a reduced delay between GFS outputs availability and the analogue forecast.

@@

The Viewer module allows displaying the resulting forecast files in an interactive GIS environment (Fig. \ref{figure:atmoswing-viewer-gui}). During the forecast, different parametrization variants of the same type of AM may be applied to a region; the reason being that some parameterizations may be specific for a subregion, when its meteorological influences differ, which is often the case even in relatively small catchments \citep{Horton2012}. The Viewer automatically gathers the similar AMs and provides a composite view of the optimal forecasts per subregion. The user can however choose to display a specific parameterization for the whole region, which is more consistent in terms of homogeneity of the analogue dates.

@@

By clicking on a station (forecast points) on the map (or by selection in a dropdown list on the left), a new window appears with a plot of the forecasted time series (Fig. \ref{figure:atmoswing-viewer-timeseries}). By default, the plot contains the usual 3 considered percentiles (90$^{th}$, 60$^{th}$, and 20$^{th}$), along with the 10 best analogues with a color code from yellow (tenth) to red (first). The 10 year return period is also displayed in order to put the forecast in perspective. The user can choose to hide any data or to display supplementary information (all analogues, all 10$^{th}$ percentiles, or all return periods) in the left panel. Traces of previous days forecasts are also automatically loaded and displayed in order to inform on the stability of the forecasts.

The user can even go into further details and display the predictand cumulative distribution for a given lead time (Fig. \ref{figure:atmoswing-viewer-distribution}). This is useful when working on extreme precipitation, as comparing the distribution of all analogues versus the 10 best analogues provides information in order to interpret the forecast (trend to under/overestimation). Distribution of the analogy criteria (not shown) can also be displayed in order to identify eventual discontinuities in the criteria values.

Finally, one can see the details of the analogue dates, along with the predictand and the criteria values in an interactive spreadsheet (not shown).

@@

\citet{Djerboua2001} applied the AM developed by \citet{Guilbaud1997} in operational forecasting as part of the MAP project \citep[\textit{Mesoscale Alpine Programme}, see][]{Binder1996} during the special observation period \citep{Bougeault2001}. He noted that the 60$^{th}$ percentile was better to forecast precipitation amounts for common situations, but for strong to extreme events, the 90$^{th}$ percentile turned out to be a better indicator. It is therefore necessary to pay particular attention to the 90$^{th}$ percentile which, when reaching high values, may be indicative of a situation related to extreme precipitations, due the presence of several analogue dates with significant observed precipitations amounts in the distribution \citep{Djerboua2001}. \citet{Bontron2004} made the same observation: he found the 60$^{th}$ percentile to be best for occurrence forecasting, but the 90$^{th}$ percentile to be the most informative in order to forecast medium to high precipitation amounts ($P > P10/10$, one tenth of the 10~year return period value). In this regard, operators should pay particular attention to the distribution of the best analogues. If, for example, it is shifted to higher values, there is a risk of under-estimating the event when considering the middle of the full distribution. Obviously, if the target situation has its best analogues with high precipitations, then this situation is likely to be rare. So, when collecting a fixed number of analogue situations, the worst ones are likely to be more common and will contribute to lower the resulting precipitation distribution, therefore a trend towards underestimation.

@@

The statistical relationship is established on a calibration period that is as long as possible. For every day of this period, a search for analogues is performed, the predictand data are associated with the corresponding dates and a forecast score is calculated. During the search for analogues situations, 120 days around the target date are excluded (thus excluding data in the same year) in order to consider only truly independent candidates days.

@@

The difference of the continuous deterministic forecasts compared to discrete ones is that they provide non-discrete values that must be evaluated with distance measures. In our application, the distribution of the analogues is summarized by a chosen percentile. Available scores are the following:

@@

Most applications mainly rely on the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows assessing the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the forecast, the smaller the score. The mean CRPS of a forecast series of length $n$ can be written:

@@

\begin{enumerate}
	\item Manual selection of the following parameters:
	\begin{enumerate}
		\item meteorological variable (e.g. geopotential, temperature, relative humidity, etc.),
		\item pressure level (e.g. 700 hPa, ...),
		\item temporal window (hour of observation \textendash e.g. 6, or 12~h~UTC),
		\item initial analogue numbers (e.g. $N_{1}=50$).
	\end{enumerate}

>>>>

\title{AtmoSwing (v1.4): Analog Technique Model for Statistical weather forecastING}

@@

The Analog Method allows forecasting local meteorological variables of interest (predictand), such as the daily precipitation, on the basis of synoptic variables (predictors). It relies on the fact that similar meteorological influences are likely to result in similar local effects. This statistical relationship is thus based on archives of observed data, and used in order to operationally forecast the coming days, or to evaluate future conditions under a changing climate.

AtmoSwing is an open source software that implements the Analog Method in a very flexible way, so that different variants can be handled dynamically, by parameterization through XML files. It is made of 3 tools: the Forecaster to perform operational forecasts, the Viewer to display the results, and the Optimizer to establish the relationship between the predictand and predictors.

@@

Usually, the predictand values could be derived in a rather physical-dynamic approach  by modeling the chain of processes linking the predictors to the predictand. The processes involved range from large scale dynamical states of the atmosphere down to very small scale microphysical processes, requiring models that are extremely complex, data-demanding, and time consuming. Conversely, given an appropriate set of predictors archives, enough situations analogous to a target one could be found so that reasonable values will be obtained for the predictand, at a reasonable coding and computing-time costs. This is particularly true for a predictand much required in hydrometeorological applications, namely the precipitation amount over a given domain and time duration. Incidentally, the forecast will be proposed as a statistical distribution based on the values taken by the predictand in the set of analogs selected, unless one considers only the single best analog, which may not prove the most efficient \citep{Bontron2005}.

In addition to precipitation forecasting, the Analog Method (AM), or an equivalent, was also used for short to medium term forecasting of daily temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983}, wind \citep{Gordon1987}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}. \citet{Guilbaud1997} performed a literature review about the use of the AM in long-term forecasting and identified operational applications for monthly forecasts in many countries, including Canada \citep{Shabbar1986},  Hungary \citep{Toth1989}, the Netherlands \citep{Nap1981}, and England \citep{Murray1974}, as well as seasonal forecasts: \citet{Barnett1978}, \citet{Bergen1982} and \citet{Livezey1988}.

@@

\section{The Analog Method}
\label{sec:method}

@@

The AM is based on the principle that two similar synoptic situations may produce similar local effects \citep{Lorenz1956}. The perfect analogy does not exist, but sufficiently similar situations leading to similar effects can be identified. Thus, two states of the atmosphere that are alike are called analogs \citep{Lorenz1969}. To be relevant, this analogy must be selected by optimizing the following elements:

\begin{itemize}
	\item The analogy variable (the predictor) must contain synoptic information having indirect dependency with the local weather and the target predictand (often the precipitation).
	\item The pressure level, or height, at which the analogy variable is considered.
	\item The spatial window is the domain on which the analogy variable is considered. The ideal size of this area is the one that maximizes useful information.
	\item The temporal window is the hour in the day, or a set of different hours, at which the analogy variables are considered.
	\item The preselection, which is often a period of the year in which similar situations are being sought. Thus, only situations with a similar distribution of solar energy are compared \citep{Lorenz1969}.
	\item The analogy criterion, needed to compare the variables on the chosen spatial and temporal windows, is a score used for ranking past situations according to their degree of similarity with the target situation.
	\item The ideal number of analog situations is the best compromise in order to take into account local variability and to maximize useful synoptic information \citep{Bontron2004}.
\end{itemize}

Because of the chaotic nature of the atmosphere, two analog situations quickly diverge over time \citep{Lorenz1969}. Thus, the AM has strong limitations regarding the temporal extrapolation \citep{Bontron2004}. Numerical models being more capable of simulating the dynamic evolution of the atmosphere, the temporal extrapolation of the synoptic variables is left to them. The search for analogy aims thus at connecting the forecasted synoptic situation with a local predictand (temperature, precipitation, etc), which is more difficult to simulate by numerical models.

@@

	The $N_{1}$ dates with the lowest values of S1 are considered as analogs, in terms of the general circulation, to the target day. The number of analogs, $N_{1}$, is a parameter to calibrate.

	\item Subsequent level(s) of analogy: beyond the similarity of airflows, one may look for similarities in other variables of interest, such as moisture variables. Therefore, the $N_{1}$ analogs are subsampled once again on the basis of another variable in order to obtain a lower number of analog dates, $N_{2}$. The S1 criterion is usually not relevant for other predictors that the atmospheric circulation. Classic criteria represent absolute distances: Mean Absolute Error (MAE) and Root Mean Squared Error (RMSE), the latest being most often used.

	This process can be repeated, by subsampling a decreasing number of analogs, $N_{i}$, according to various meteorological variables.

@@

The basis of the method was introduced by \citet{Lorenz1969}, who considered the analogy of the geopotential field on different pressure levels. In addition, the analog situations were to belong to the same period of the year to be comparable in terms of distribution of solar energy. This preselection was later implemented as a moving selection of more or less two months around the target date.

@@

Every variable of the NCEP/NCAR reanalysis was systematically assessed for each of the identified subregions of the study area (Sect. \ref{sec:case_study}). The present assessment considers every subregion separately and optimizes the spatial window and the number of analogs for every combination of variables, pressure levels, times of observation, and subregions (using the sequential calibration tool implemented in AtmoSwing, see Sect. \ref{sec:atmoswing-calibration}). Even though the best predictors may vary from one place to another due to different meteorological influences (see Fig. \ref{figure:variable_exploration_Chablais} for the 30 best predictors for the Chablais region, and Fig. \ref{figure:variable_exploration_SE_crests} for the Southeast ridges), 2 main conclusions can be drawn from the NCEP/NCAR reanalysis dataset for the area of interest, that meet those from \citet{Bontron2004}: (1) the variables describing the atmospheric circulation (pressure fields or geopotential heights) are the ones with the highest prediction capacity (the precipitation rate is even rarely selected in the 30 best predictors), and (2) these are better when compared by means of the S1 criteria (asterisk in Fig. \ref{figure:variable_exploration_Chablais} and \ref{figure:variable_exploration_SE_crests}) instead of the RMSE. The reason for this latter point is the better representation of the atmospheric circulation information when considering the shape of a pressure field rather than its values.

\citet{Horton2012} identified that the analogy on the atmospheric circulation as defined by \citet{Bontron2004} may not be comprehensive in an alpine environment, and that considering 4 atmospheric levels (1000, 850, 700, and 500~hPa at 12 or 24~h) instead of 2 could improve the performance (Table \ref{table:R6} -- the number of analogs $N_{1}$ was not provided globally, as it was calibrated for every subregion). The change in performance scores is illustrated in Fig. \ref{figure:scores_1} for the different subregions of the study area. An equivalent structure was also optimized by means of generic algorithms (PC-4Zgo), which is the topic of another paper \citep{Horton2016b}. As these new parametrizations provide performance gains for alpine catchments under varying meteorological influences (Fig. \ref{figure:scores_1}), they are recommended for operational use.

@@

\citet{Horton2012} also integrated an analogy on the moisture flux, after the circulation analogy on the 4 atmospheric levels (Table \ref{table:R7}). The number of analogs ($N_{1}$, $N_{2}$) were calibrated for every subregion. Similarly, an equivalent structure was optimized \citep[PC-4Zgo-2MIgo,][]{Horton2016b}. The change in the performance scores, for the method with 2 levels of analogy, is illustrated in Fig. \ref{figure:scores_2} for the study area. As previously, we would recommend using these new parametrizations for alpine catchments.

@@

\citet{Hamill2006} used an analogy based approach on the GFS reforecasts in order to correct systematic errors in the ensemble forecasts of temperature and precipitation. Indeed, the statistical approach pointed at some bias in estimates of the numerical model, which may have been adjusted by taking into account the intrinsic local climatology from the AM. Moreover, the under-dispersion of the ensemble forecast from the numerical model has been corrected using analogs \citep{Hamill2006}. Correction of ensemble forecast under-dispersion by means of the AM is also used operationally at EDF (\'{E}lectricit\'{e} de France).

@@

Another limitation is the fact that extreme events may be under-represented in the considered sample of analog dates. Indeed, in a limited weather archive, events with high return periods are not so numerous. Their number is certainly lower than the standard number of analogs considered, which can introduce a bias in the forecast. There are however techniques to correct this bias \citep[see][]{Marty2010}.

@@

The source code in under version control (with Mercurial), and is open source (on Bitbucket, www.atmoswing.org). Developments have been made with a test driven development (TDD) approach. A collection of more than 500 tests are frequently evaluated and completed during development, in order to verify that code changes do not introduce regression. Every analogy criterion, prediction score, searching and sorting functions, data manipulation, and so on, is tested. The integration tests specific to the AM rely on the results of another analog sorting software developed at the LTHE \textendash INPG university of Grenoble. They ensure that the results of AtmoSwing are exactly equivalent to their model, given the same parameters and data. These tests are part of the checks carried out regularly.

@@

An AtmoSwing great strength is that it is designed to process analogy forecasting in a modular fashion. The structure of the AM (number of analogy levels, number of predictors) is built dynamically (Fig. \ref{figure:flowchart_modules_atmoswing}), and nothing is fixed a priori. The software then performs successively as many analogy levels as the user specified, with all the predictors he wants. Each level of analogy processed results in an object containing target dates, analog dates, values of the analogy criteria, and other data. This object can be saved in a NetCDF file and/or can be injected into a new analogy level. The whole structure of the AM is defined in an XML file generated by the user. Even the time step of the method (6 or 24~hours for example) is a dynamic parameter.

@@

Although processing an analog forecast for a given target date is fast, when optimizing the AM, forecasts are re-processed over long periods of several decades, which may become very time consuming. Thus, great effort has been made in order to reduce the processing time to a minimum. Waste of time were identified with profiling tools and cut down at different levels. The software also use the linear algebra library Eigen 3 \citep{Guennebaud2010} for calculations on vectors and matrices, which resulted in time saving. Multi-threading is also implemented so that the search for analog situations in the archive is distributed among the available threads.

The code optimization consited in reducing the most significant processing time identified by means of profiling tools. First, every identified redundancy in processing has been removed. Then, when looking for a certain date or data, the search looks first in the proximity where it is likely to be found instead of exploring an entire array. Similar data are not loaded twice, but shared pointers are used. Many other improvements allowed saving more time, for example the use of the quicksort method \citep{Hoare1962a} to sort the date vectors according to the values of analogy criteria. Different implementations were tested (many remain in the code as alternatives) in order to select the most efficient: for example, when storing analog dates according to their criteria value, it is faster to insert them in a fixed-size array rather than storing them all and sorting the array subsequently.

The major performance gains were obtained by:
\vspace*{-2mm}
\begin{itemize}
	\setlength\itemsep{-2px}
	\item reduction of the array sizes used in the calculations,
	\item use of pointers (and therefore the decrease of data copies),
	\item better search for dates in temporal vectors starting from the previous index (reduction of the range of the array to explore),
	\item better management of the analog vectors,
	\item preprocessed gradients on geopotential fields when using the S1 criteria,
	\item and finally the implementation of parallel calculations (multi-threading).
\end{itemize}

@@

The software first downloads the predictor describing the target situation, such as 1\textdegree\ GFS outputs \citep[Global Forecast System,][see Sect. \ref{sec:data}]{Kanamitsu1991,Kanamitsu1989}. It then interpolates linearly the gridded data to match the resolution of the archive (for example 2.5\textdegree\ in the case of the NCEP/NCAP reanalysis I). The analogs dates are next sought according to the selected parameterization, and the predictand data are associated with the corresponding dates. The results are finally saved in auto-describing NetCDF files. If required, a synthetic XML is generated for an easier integration on a web platform for example. Every step of the forecast, from the predictor downloading to the final results, is done in the software (and controlled through configuration), without use of external scripts (e.g. for data conversion).

When using a given parameterization of the AM, it has previously been optimized in a perfect prognosis context \citep[i.e. by optimizing on the reanalysis dataset; however, this does not take into account the uncertainty of the target situation when taken from operational forecasts of a Numerical Weather Prediction model,][]{Klein1963}. In operational forecasting, one should take into account the uncertainty of the predictor increasing with the lead time. A solution, presented in Sect. \ref{sec:optimization-operational}, consists in increasing the number of analog situations to keep with the lead time. This technique is available in AtmoSwing, as the number of analogs can be specified for every lead time, making such approach easily accessible.

The command-line interface allows to start the process for the present date, a selected past date, or the last $x$ days (as long as the predictors are available on the provider server). When there is nothing to process (no new predictor data available), the execution just stops, leading to no loss of computing resources. The recommended use is thus to set up a cron task on a Linux server in order to trigger the forecast every 30 minutes. When using GFS outputs, this would provide 4 forecasts a day with a reduced delay between GFS outputs availability and the analog forecast.

@@

The Viewer module allows displaying the resulting forecast files in an interactive GIS environment (Fig. \ref{figure:atmoswing-viewer-gui}). During the forecast, different parametrization variants of the same type of AM may be applied to a region; the reason being that some parameterizations may be specific for a subregion, when its meteorological influences differ, which is often the case even in relatively small catchments \citep{Horton2012}. The Viewer automatically gathers the similar AMs and provides a composite view of the optimal forecasts per subregion. The user can however choose to display a specific parameterization for the whole region, which is more consistent in terms of homogeneity of the analog dates.

@@

By clicking on a station (forecast points) on the map (or by selection in a dropdown list on the left), a new window appears with a plot of the forecasted time series (Fig. \ref{figure:atmoswing-viewer-timeseries}). By default, the plot contains the usual 3 considered percentiles (90$^{th}$, 60$^{th}$, and 20$^{th}$), along with the 10 best analogs with a color code from yellow (tenth) to red (first). The 10 year return period is also displayed in order to put the forecast in perspective. The user can choose to hide any data or to display supplementary information (all analogs, all 10$^{th}$ percentiles, or all return periods) in the left panel. Traces of previous days forecasts are also automatically loaded and displayed in order to inform on the stability of the forecasts.

The user can even go into further details and display the predictand cumulative distribution for a given lead time (Fig. \ref{figure:atmoswing-viewer-distribution}). This is useful when working on extreme precipitation, as comparing the distribution of all analogs versus the 10 best analogs provides information in order to interpret the forecast (trend to under/overestimation). Distribution of the analogy criteria (not shown) can also be displayed in order to identify eventual discontinuities in the criteria values.

Finally, one can see the details of the analog dates, along with the predictand and the criteria values in an interactive spreadsheet (not shown).

@@

\citet{Djerboua2001} applied the AM developed by \citet{Guilbaud1997} in operational forecasting as part of the MAP project \citep[\textit{Mesoscale Alpine Programme}, see][]{Binder1996} during the special observation period \citep{Bougeault2001}. He noted that the 60$^{th}$ percentile was better to forecast precipitation amounts for common situations, but for strong to extreme events, the 90$^{th}$ percentile turned out to be a better indicator. It is therefore necessary to pay particular attention to the 90$^{th}$ percentile which, when reaching high values, may be indicative of a situation related to extreme precipitations, due the presence of several analog dates with significant observed precipitations amounts in the distribution \citep{Djerboua2001}. \citet{Bontron2004} made the same observation: he found the 60$^{th}$ percentile to be best for occurrence forecasting, but the 90$^{th}$ percentile to be the most informative in order to forecast medium to high precipitation amounts ($P > P10/10$, one tenth of the 10~year return period value). In this regard, operators should pay particular attention to the distribution of the best analogs. If, for example, it is shifted to higher values, there is a risk of under-estimating the event when considering the middle of the full distribution. Obviously, if the target situation has its best analogs with high precipitations, then this situation is likely to be rare. So, when collecting a fixed number of analog situations, the worst ones are likely to be more common and will contribute to lower the resulting precipitation distribution, therefore a trend towards underestimation.

@@

The statistical relationship is established on a calibration period that is as long as possible. For every day of this period, a search for analogs is performed, the predictand data are associated with the corresponding dates and a forecast score is calculated. During the search for analogs situations, 120 days around the target date are excluded (thus excluding data in the same year) in order to consider only truly independent candidates days.

@@

The difference of the continuous deterministic forecasts compared to discrete ones is that they provide non-discrete values that must be evaluated with distance measures. In our application, the distribution of the analogs is summarized by a chosen percentile. Available scores are the following:

@@

Most applications mainly rely on the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows assessing the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the forecast, the smaller the score. The mean CRPS of a forecast series of length $n$ can be written:

@@

\begin{enumerate}
	\item Manual selection of the following parameters:
	\begin{enumerate}
		\item meteorological variable (e.g. geopotential, temperature, relative humidity, etc.),
		\item pressure level (e.g. 700 hPa, ...),
		\item temporal window (hour of observation \textendash e.g. 6, or 12~h~UTC),
		\item initial analog numbers (e.g. $N_{1}=50$).
	\end{enumerate}

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification of the most skilled unitary cell (1~point for moisture variables and 4 for the geopotential height when using the S1 criteria) over a large domain. Every point (or cell) of the full domain is assessed jointly on all predictors of the level of analogy (consisting generally of the same variable, but on different pressure levels and at different hours).
		\item From this most skilled point, the spatial window is expanded by successive iterations in the direction of greater performance gain. The detailed stages are the followings: (i) The unitary spatial window is expanded in every 4 directions successively. The performance score is processed for these 4 windows. (ii) Only the direction providing the best improvement is applied to the spatial window. (iii) From this new spatial window, an increase in every 4 directions is once again assessed, and the best improvement is applied. (iv) The spatial window grows up by repeating the previous steps, until no improvement is reached.
		\item The number of analog situations is then reconsidered and optimized for the current level of analogy.
		\item A new level of analogy can then be added, based on other variables (such as the moisture index) on predefined pressure levels and time frames. The number of analogs for the next level of analogy is initiated at a chosen value. Then, the procedure starts again from step (a) for the new level. The parameters calibrated on the previous analog levels are fixed and do not change (except the numbers of analogs of all levels of analogy, that are optimized once again at the final stage). 
	\end{enumerate}
	\item Finally, the numbers of analogs are re-assessed for the different levels of analogy. This is done iteratively by varying the number of analogs of each level in a systematic way.
\end{enumerate}
